---
title: "Intro to Vroom"
output: html_notebook
---

```{r}
library(vroom)
library(fs)
library(purrr)
library(dplyr)
```

# Introduction to `vroom`

## `vroom` basics
*Load data into R using `vroom`*

1. Load the `vroom()` library
  ```{r}
  library(vroom)
  ```

1. Use the `vroom()` function to read the **transactions_1.csv** file from the **../data/** folder
```{r}
vroom("../data/transactions_1.csv")
```

1. Use the `id` argument to add the file name to the data frame.  Use **file_name** as the argument's value
```{r}
vroom("../data/transactions_1.csv", id = "file_name")
```

1. Load the prior command into a variable called `vr_transactions`
```{r}
vr_transactions <- vroom("../data/transactions_1.csv", id = "file_name")

vr_transactions
```

1. Load the file spec into a variable called `vr_spec`, using the `spec()` command
```{r}
vr_spec <- spec(vr_transactions)

vr_spec
```

## Load multiple files

1. Load the `fs` and `dplyr` libraries
```{r}
library(fs)
library(dplyr)
```

1. List files in the **../data/** folder using the `dir_ls()` function
```{r}
 dir_ls("../data")
```

1. In the `dir_ls()` function, use the `glob` argument to pass a wildcard to list CSV files only. Load to a variable named `files`
```{r}
files <- dir_ls("../data", glob = "*.csv")
```

1. Pass the `files` variable to `vroom`. Set the `n_max` argument to 1,000 to limit the data load for now
```{r}
vroom(files, n_max = 1000)
```

1. Add a `col_types` argument with `vr_specs` as its value
```{r}
vroom(files, n_max = 1000, col_types = vr_spec)
```

1. Use the `col_select` argument to pass a `list` object containing the following variables: order_id, date, customer_name, and price
```{r}
vroom(files, n_max = 1000, col_types = vr_spec,
      col_select = list(order_id, date, customer_name, price)
      )
```

## Load and modify multiple files
*For files that are too large to have in memory, keep a summarization*

1. Use a `for()` loop to print the content of each vector inside `files`
```{r}
for(i in seq_along(files)) {
  print(files[i])
}
```

1. Switch the `print()` command with the `vroom` command, using the same arguments, except the file name.  Use the `files` variable.  Load the results into a variable called `transactions`. 
```{r}
for(i in seq_along(files)) {
  transactions <- vroom(files[i], n_max = 1000, col_types = vr_spec, 
                       col_select = list(order_id, date, customer_name, price))

}
```

1. Group `transactions` by `order_id` and get the total of `price` and the number of records. Name them `total_sales` and `no_items` respectively. Name the new variable `orders`
```{r}
for(i in seq_along(files)) {
  transactions <- vroom(files[i], n_max = 1000, col_types = vr_spec, 
                       col_select = list(order_id, date, customer_name, price))
  orders <- transactions %>%
    group_by(order_id) %>%
    summarise(total_sales = sum(price), no_items = n())
}
```

1. Define the `orders` variable as `NULL` prior to the for loop and add a `bind_rows()` step to `orders` to preserve each summarized view.
```{r}
orders <- NULL
for(i in seq_along(files)) {
  transactions <- vroom(files[i], n_max = 1000, col_types = vr_spec, 
                       col_select = list(order_id, date, customer_name, price))
  orders <- transactions %>%
    group_by(order_id) %>%
    summarise(total_sales = sum(price), no_items = n()) %>%
    bind_rows(orders)
}
```
    
1. Remove the `transactions` variable at the end of each cycle    
```{r}
orders <- NULL
for(i in seq_along(files)) {
  transactions <- vroom(files[i], n_max = 1000, col_types = vr_spec, 
                       col_select = list(order_id, date, customer_name, price))
  orders <- transactions %>%
    group_by(order_id) %>%
    summarise(total_sales = sum(price), no_items = n()) %>%
    bind_rows(orders)
  rm(transactions)
}
```
    
1. Preview the `orders` variable
```{r}
orders
```

